---
title: "harmonic analysis"
author: "time series group 4"
date: "2023-02-23"
output: html_document
---

Since the p-value 0.1>0.05, we can accept the null hypothesis that the differenced time series is stationary, excluding the seasonality. 

```{r}
kpss.test(BoxCox(train_ord, lamo)) # is stationary 
```
our data is stationary


```{r}
BoxCox(train_ord, lamo)%>% mstl() %>% autoplot()
```
See that our seasonality is not sufficient

#Note, this is the plot we should use, refer to the data below to label it 

```{r}
periodogram(BoxCox(train_ord, lamo),log='no',plot=TRUE,ylab="Periodogram",xlab="Frequency")
```
Using the periodogram, we can see which spikes are largest 

We need help identifying the frequency of these spikes

```{r}
temp<-periodogram(BoxCox(train_ord, lamo),log='no',plot=FALSE,ylab="Periodogram",xlab="Frequency",xlim=c(0,0.1))

frequencies=data.frame(freq=temp$freq, spec=temp$spec)
frequencies=frequencies[order(-frequencies$spec),]
head(frequencies,20)
```
The frequencies drop off after the first two. hopefully these frequencies will be sufficient 

```{r}
ggplot(frequencies, aes(x = freq, y = spec)) + 
  geom_point() +
  geom_vline(aes(xintercept = frequencies[1,"freq"], color = "1st"), linetype = "dashed")+
    geom_vline(aes(xintercept = frequencies[2,"freq"], color = "2nd"), linetype = "dashed")+
    geom_vline(aes(xintercept = frequencies[3,"freq"], color = "3rd"), linetype = "dashed")+
    geom_vline(aes(xintercept = frequencies[4,"freq"], color = "4th"), linetype = "dashed")+
    geom_vline(aes(xintercept = frequencies[5,"freq"], color = "5th"), linetype = "dashed")+
  xlab("Frequency") +
  ylab("Spectrum")+ 
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        legend.key = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, face = "bold"))

```

#interpret frequencies

The relevant frequencies are as follows: 
```{r}
freq=1/frequencies[1:5,"freq"]
freq
```
Surprisingly 53 is not appearing as a suggested period unil number 5

We are getting 106, which is periods of 2 years, and 26 which is periods of half a year. The next frequencies had lower spectrum values: 
6.666 (e.g 46.6 days), which is a month and a half. 
53.33, which is one year
160, which is three years. 


```{r}
train_ord %>% msts(ts.frequency=53,seasonal.periods = freq[1],start=c(2013,1)) %>% mstl(lambda=lamo)%>%autoplot()
train_ord %>% msts(ts.frequency=53,seasonal.periods = freq[1:2],start=c(2013,1)) %>% mstl(lambda=lamo)%>%autoplot()

```

Using one period is sufficient, while two periods are not

# evaluate various k values 

```{r}
freq2=append(freq[1:2],53)

msts_2 <- train_ord%>% msts(ts.frequency=53,seasonal.periods = freq2[1:3] ,start=c(2013,1)) #freq[1:3]
list_k1 <- c()
list_k2 <- c()
list_k3 <- c()
list_lb <- c()
list_rmse <- c()

                   
for (k1 in 1:6){
  for (k2 in 0:(7-k1)){
      k3=0        
      print(c(k1,k2,k3))
      #if k2>4
      fit_2 <- auto.arima(msts_2, lambda='auto', xreg=fourier(msts_2,K=c(k1,k2,k3)), seasonal=FALSE)
      list_k1 = append(list_k1,k1)
      list_k2 = append(list_k2,k2)
      list_k3 = append(list_k3,k3)
      list_lb = append(list_lb,Box.test(fit_2$residuals, type='Ljung-Box', lag=53)$p.value)
      rmse=sum((fit_2$residuals)^2)
      list_rmse = append(list_rmse,rmse)
    }
  }




results=data.frame(
              k1=list_k1,
              k2=list_k2,
              lb_pval=list_lb,
              rmse_val=list_rmse
              )

results$ss=results$lb_pval>0.05

filter(results,ss)

#1,1,0 is the most minimal

#i like 4,0 , but 4,1 could work 

```

comparing strong values: 
```{r}
#here are the values to itterate through 
k1_final=c(1,1,2,3)
k2_final=c(1,5,1,1)

#redo
k1_final=c(1,2,3,4,4,4,5,6)
k2_final=c(0,0,0,0,1,2,1,1)


for (i in 1:8){
  fit <- auto.arima(msts_2, lambda='auto', xreg=fourier(msts_2,K=c(k1_final[i],k2_final[i],0)), seasonal=FALSE)
  #checkresiduals(fit)
  #print(fit$aicc) #AICC is basically the same for all of them
  
  
  fc<-forecast(fit,xreg=fourier(msts_2,K=c(k1_final[i],k2_final[i],0),53))
  #plot(fc)
  print("------------------------------------------------------------------------------")
  print(paste0("Model fourier(",k1_final[i],",",k2_final[i],",0)"))
  print(accuracy(fc$mean, test_ord))
}

#lowest AICC from the third model
#pick c(1,0,0), c(4,1,0)

```
K=c(1,1,0) is the best because it has the lowest RMSE and AICC


Testing it out
```{r}
fit_2 <- auto.arima(msts_2, lambda=lamo, xreg=fourier(msts_2,K=c(4,1,0)), seasonal=FALSE)
checkresiduals(fit_2)
#Arima(msts_2,order=c(1,0,0),lambda=lamo, xreg=fourier(msts_2,K=c(1,1,0)))%>%checkresiduals()
Box.test(fit_2$residuals, type='Ljung-Box', lag=53)$p.value

print(fit_2$aicc)

fit_3 <- auto.arima(msts_2, lambda=lamo, xreg=fourier(msts_2,K=c(1,0,0)), seasonal=FALSE)
checkresiduals(fit_3)
#Arima(msts_2,order=c(1,0,0),lambda=lamo, xreg=fourier(msts_2,K=c(1,1,0)))%>%checkresiduals()
Box.test(fit_3$residuals, type='Ljung-Box', lag=53)$p.value

print(fit_3$aicc)
```
ARIMA(1,0,0) is sufficient

```{r}
fc_msts<-forecast(fit_2,xreg=fourier(msts_2,K=c(4,1,0),53))
plot(fc_msts)
accuracy(fc_msts$mean, test_ord)


fc_msts<-forecast(fit_3,xreg=fourier(msts_2,K=c(1,0,0),53))
plot(fc_msts)
accuracy(fc_msts$mean, test_ord)

```

Also try the TBATS method 


```{r}
tbats_fit = tbats(train_ord )
checkresiduals(tbats_fit)

#white noise
Box.test(tbats_fit$errors, type='Ljung-Box', lag=53)$p.value

fc_tbats=forecast(tbats_fit,53)
autoplot(fc_tbats)

accuracy(fc_tbats$mean, test_ord)

```
Same box cox, yearly seasonality, and 8 week seasonality 
Unfortunately the error is not white noise

```{r}

autoplot(fc_tbats)
autoplot(fc_msts)

# ggplot(ord2, aes(x))
#   geom_line(aes(y = ARIMA_exp, color = "ARIMA Exp")) +
#   geom_line(aes(y = ARIMA_slide, color = "ARIMA Slide")) +
#   geom_line(aes(y = ETS_exp, color = "ETS Exp")) +
#   geom_line(aes(y = ETS_slide, color = "ETS Slide")) +
#   scale_color_manual(values = c("ARIMA Exp" = "blue", "ARIMA Slide" = "blue4", "ETS Exp" = "coral", "ETS Slide" = "coral4")) +
#   xlab("iteration number") +
#   ylab("AICc") +
#   ggtitle("AICc by Model") +
#   theme(plot.title = element_text(hjust = 0.5)) +
#   theme(legend.title = element_blank(),
#         legend.position = "bottom")

```



