---
title: "final_project"
author: "Maggie Chen"
date: "2023-02-12"
output: html_document
---

```{r}
library(tseries)
library(forecast)
library(dplyr)
library(reshape2)
library(tidyverse)
library(tidyr)
```

We have flight data to and from ORD to LAX from 2013 to 2022. 
Cancellation Codes: 
A: carrier
B: weather
C: national air system
D: security


#### Read ORD origin and LAX origin files
1. read the file and convert NA to 0
```{r}
setwd("/Users/eujeneyum/Desktop/UChicago_Courses/Q5/time_series/final_project/")
ord_flights <- read.csv('ord_origin_master.csv')
ord_flights[is.na(ord_flights)] <- 0
ord_flights[ord_flights == ""] <- "N/A"
lax_flights <- read.csv('lax_origin_master.csv')
lax_flights[is.na(lax_flights)] <- 0
lax_flights[lax_flights == ""] <- "N/A"
```



```{r}
ord_flights <- ord_flights %>% mutate(row_id = row_number())
ord_flights <- ord_flights %>% pivot_wider(names_from = CANCELLATION_CODE, values_from = FLIGHTS, values_fn = length)
ord_flights [is.na(ord_flights)] <- 0
ord_flights <- select(ord_flights, -c("N/A"))
```


```{r}
lax_flights <- lax_flights %>% mutate(row_id = row_number())
lax_flights <- lax_flights %>% pivot_wider(names_from = CANCELLATION_CODE, values_from = FLIGHTS, values_fn = length)
lax_flights [is.na(lax_flights)] <- 0
lax_flights <- select(lax_flights, -c("N/A"))
```


```{r}

```


2.1 aggregate the dataset for ORD & LAX
```{r}
df_ord = ord_flights %>% 
                  group_by(FL_DATE, OP_UNIQUE_CARRIER) %>%
                  summarize(
                    arr_tot_delay = sum(ARR_DELAY_NEW),
                    arr_num_delay = length(ARR_DELAY_NEW),
                    arr_avg_delay = mean(ARR_DELAY_NEW),
                    cancel_A = sum(A),
                    cancel_B = sum(B),
                    cancel_C = sum(C),
                    cancel_D = sum(D),
                    delay_weather = sum(WEATHER_DELAY),
                    delay_carrier = sum(CARRIER_DELAY),
                    delay_nas = sum(NAS_DELAY),
                    delay_security = sum(SECURITY_DELAY),
                    delay_late_aircraft = sum(LATE_AIRCRAFT_DELAY)
                  )

```
```{r}
df_lax = ord_flights %>% 
                  group_by(FL_DATE, OP_UNIQUE_CARRIER) %>%
                  summarize(
                    arr_tot_delay = sum(ARR_DELAY_NEW),
                    arr_num_delay = length(ARR_DELAY_NEW),
                    arr_avg_delay = mean(ARR_DELAY_NEW),
                    cancel_A = sum(A),
                    cancel_B = sum(B),
                    cancel_C = sum(C),
                    cancel_D = sum(D)
                  )
```


```{r}
df_lax = lax_flights %>%
                  group_by(FL_DATE) %>%
                  summarize(
                    lax_arr_tot_delay = sum(ARR_DELAY_NEW),
                    lax_arr_num_delay = length(ARR_DELAY_NEW),
                    lax_arr_avg_delay = mean(ARR_DELAY_NEW)
                  )
```

```{r}
df = left_join(df_ord, df_lax, by=c("FL_DATE"))
```

2.2 convert to ts object and plot
```{r}
ord_arr_ts <- ts(ord_arr_delay$x, start = c(2013, 1), frequency = 365)
tsdisplay(ord_arr_ts)
```
3.1 aggregate the sum of ARR_DELAY_NEW based on months
```{r}
ord_flights$month <- format(as.Date(ord_flights$DATE, format="%d/%m/%Y"),"%m")
ord_arr_delay <- aggregate(ARR_DELAY_NEW ~ month + YEAR, ord_flights, FUN = sum )
```

3.2 convert to ts object and plot
```{r}
ord_arr_ts <- ts(ord_arr_delay$ARR_DELAY_NEW, start = c(2013, 1), frequency = 12)
tsdisplay(ord_arr_ts)
```
questions:
1. how to aggregate - sum or avg for a single day?
2. the scope - do we want to narrow down to one way & maybe a single airline?
3. frequency of ts object - do we want to do daily or monthly?



#### TESTING WITH JUST SMALL DATA
BRING IN DATA
```{r}
flights <- read.csv('flight_delays_2015_LAX_ORD_NDLAX.csv')
flights[is.na(flights)] <- 0
flights[flights == ""] <- "N/A"
```


```{r}
flights$flights_uid = paste0(flights$FL_DATE, "_", flights$OP_CARRIER_FL_NUM, "_", flights$OP_UNIQUE_CARRIER)
flights <- distinct(flights, flights_uid, .keep_all = TRUE)
```


CREATE HOUR COLUMN
```{r}
flights <- flights %>%
  mutate(
    HOUR = substr(as.character(CRS_ARR_TIME),1,nchar(as.character(CRS_ARR_TIME))-2)
    )
```


```{r}
flights$HOUR[flights$HOUR == ""] <- '0'
flights = flights %>% 
                mutate(
                  HOUR = if_else(nchar(HOUR) == 1, paste0('0', HOUR), HOUR))
```


```{r}
flights$FLIGHTS = 1
```


```{r}
flights <- flights %>% mutate(row_id = row_number())
flights <- flights %>% pivot_wider(names_from = CANCELLATION_CODE, values_from = FLIGHTS, values_fn = length)
flights [is.na(flights)] <- 0
flights <- select(flights, -c("N/A"))
```

```{r}
df = flights %>% 
      group_by(FL_DATE, OP_UNIQUE_CARRIER, ORIGIN, HOUR) %>%
      summarize(
        arr_tot_delay = sum(ARR_DELAY_NEW),
        arr_num_delay = length(ARR_DELAY_NEW),
        arr_avg_delay = mean(ARR_DELAY_NEW),
        dep_tot_delay = sum(DEP_DELAY_NEW),
        dep_num_delay = length(DEP_DELAY_NEW),
        dep_avg_delay = mean(DEP_DELAY_NEW),
        delay_weather = sum(WEATHER_DELAY),
        delay_carrier = sum(CARRIER_DELAY),
        delay_nas = sum(NAS_DELAY),
        delay_security = sum(SECURITY_DELAY),
        delay_late_aircraft = sum(LATE_AIRCRAFT_DELAY),
        cancel_A = sum(A),
        cancel_B = sum(B),
        cancel_C = sum(C)
        #cancel_D = sum(D)
      )

# BRING BACK THE D COLUMN LATER 
#df$HOUR = as.integer((df$HOUR))

```

```{r}
df$DATE = as.Date(df$FL_DATE, format = '%m/%d/%Y')
```

```{r}
hours = unique(flights$HOUR)
hours = hours[order(hours)]
ref <- expand.grid(id = unique(flights$DATE), t = hours)
table(ref$t)
ref$tid = paste0(ref$id, " ", ref$t)
ref <- expand.grid(id = unique(ref$tid), t = unique(flights$OP_UNIQUE_CARRIER))
ref = ref[order(ref$id, ref$t),]
ref$origin = 'ORD'
ref2 <- ref
ref2$origin = "LAX"
ref_final <- rbind(ref, ref2)
colnames(ref_final) <- c('datetime','carrier', 'origin')

```



JOIN

```{r}
df$datetime = paste0(df$DATE, " ", df$HOUR)
df_final = left_join(ref_final,df,by=c("datetime"="datetime","carrier"="OP_UNIQUE_CARRIER", "origin" = "ORIGIN"))
```

```{r}
#head(df_final)
df_final$HOUR = substr(df_final$datetime, nchar(df_final$datetime)-2, nchar(df_final$datetime))
df_final$DATE = substr(df_final$datetime, 1, nchar(df_final$datetime)-3)
df_final = df_final[order(df_final$DATE, df_final$HOUR),]
df_final = select(df_final,-c(FL_DATE))
df_final[is.na(df_final)] <- 0
head(df_final)
```

TIME SERIES
```{r}
# carrier, origin
write.csv(df_final,'aggregated_data.csv')
```

```{r}

```

